Type Classes for Mathematics in Type Theory

### 2010-06-30 jar.tex
Fix two typos.

For more complex properties such as injectivity of partially applied addition and multiplication, we take a longer detour. As part of our universal algebra theory, we have proved that proofs of statements in the language of an equational theory can be transferred between isomorphic implementations. Combined with the trivially proved property that any two initial objects in a category are isomorphic, this lets us tranfer proofs of such statements between implementations of \lstinline|Naturals|, requiring only that we reflect the concrete statement (expressed in terms of the operational type classes) to a symbolic statement in the language of semirings. We intend to eventually make this reflection completely automatic using type class based quotation techniques along the lines of those described in Section~\ref{quoting}.

@@

There are really only two pending concerns that keeps us from making an unequivocal endorsement of type classes as a versatile, expressive, and elegant means of organizing proof developments. The first and lesser of the two is universe polymorphism for definitions as described in the previous section. The second is instance resolution efficiency. In more complex parts of our development we are now experiencing increasingly serious efficiency problems, despite having already made sacrifices by artifically inhibiting many natural class instances in order not to further strain instance resolution.
%While we have already found the predicate-class-based style of formalization to work very well for our development thus far, and believe it has the potential to be \emph{the} modern solution to this problem, in our work we have encountered a number of limitations in the Coq implementation of key features of the tool, that keep us from making an unconditional recommendation for universal adoption of these techniques.
Fortunately, there is plenty of potential for improvement of the current instance resolution implementation. One source is the vast literature on efficient implementation of Prolog-style resolution, which the hint-based proof search used for instance resolution greatly resembles.

>>>>

For more complex properties such as injectivity of partially applied addition and multiplication, we take a longer detour. As part of our universal algebra theory, we have proved that proofs of statements in the language of an equational theory can be transferred between isomorphic implementations. Combined with the trivially proved property that any two initial objects in a category are isomorphic, this lets us transfer proofs of such statements between implementations of \lstinline|Naturals|, requiring only that we reflect the concrete statement (expressed in terms of the operational type classes) to a symbolic statement in the language of semirings. We intend to eventually make this reflection completely automatic using type class based quotation techniques along the lines of those described in Section~\ref{quoting}.

@@

There are really only two pending concerns that keeps us from making an unequivocal endorsement of type classes as a versatile, expressive, and elegant means of organizing proof developments. The first and lesser of the two is universe polymorphism for definitions as described in the previous section. The second is instance resolution efficiency. In more complex parts of our development we are now experiencing increasingly serious efficiency problems, despite having already made sacrifices by artificially inhibiting many natural class instances in order not to further strain instance resolution.
%While we have already found the predicate-class-based style of formalization to work very well for our development thus far, and believe it has the potential to be \emph{the} modern solution to this problem, in our work we have encountered a number of limitations in the Coq implementation of key features of the tool, that keep us from making an unconditional recommendation for universal adoption of these techniques.
Fortunately, there is plenty of potential for improvement of the current instance resolution implementation. One source is the vast literature on efficient implementation of Prolog-style resolution, which the hint-based proof search used for instance resolution greatly resembles.
